---
title: "Oct 25 Notes"
author: "Hongdou Li"
date: "10/25/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Recap:

+ strict vs. weak stationary.

+ Mean function $\mu(t)=E[Y_t]$

+ Covariance function: $\gamma(t,t+h)=Cov(Y_t,Y_{t+h})$

In the case that $Y_t$ is a stationary time series, $\gamma(t,t+h)=\gamma(h)$ is called the **autocovariance function**.

In this context, we define the **autocorrelation function** of lag h to be 

$\rho(h)=Corr(Y_t,Y_{t+h})$

>> $=\frac{Cov(Y_t,Y_{t+h})}{\sqrt{Var(Y_t)Var(Y_{t+h})}}$.

>> $=\frac{Cov(Y_t,Y_{t+h})}{\sqrt{Cov(Y_t,Y_{t})Cov(Y_{t+h},Y_{t+h})}}$

>> $\frac{\gamma(h)}{\gamma(0)}$

* Properties:

+ $\gamma(0)\geq0$ <=> $Var[Y_t]\geq0$

+ $|\gamma(h)|\leq\gamma(0)$ <=> $\frac{|\gamma(h)|}{\gamma(0)}=|\rho(h)|\leq1$

+ $\gamma(h)=\gamma(-h)$, $\rho(h)=\rho(-h)$ (just need to care about the positive side)

## Example1:

Recall that if ${Y_t}\sim MA(1)$,

$\gamma(h)=Cov(Y_t,Y_{t+h})=\begin{cases}\sigma^2(1+\theta^2)&h=0\\\sigma^2\theta&h=\pm1\\0& otherwise\end{cases}$ 

Notice $\gamma(0) = \sigma^2(1+\theta^2)$ and so 

$\rho(h)=\frac{\gamma(h)}{\gamma(0)}=\begin{cases}1&h=0\\\theta/(1+\theta^2)&h=\pm1\\0&otherwise\end{cases}$

we can make the ACF plot based on it.

## Example2:
First order Autoregressive Model[AR(1)]. Assume ${Y_t}$ is a stationary times series satisfying the following coditions

\begin{center}
$Y_t=\phi Y_{t-1} +\epsilon_t$
\end{center}

where $|\phi|<1$ (we require this for it to be stationary, this is called the stationary condition) and ${\epsilon_t}\sim WN(0,\sigma^2)$ and $\epsilon_t$ and $Y_s$ are uncorrelated for s<t. Calculate $\gamma(h)$ and $\rho(h)$.

+ $E[Y_t]=E[\phi Y_{t-1} +\epsilon_t]=\phi E[ Y_{t-1}] +E[\epsilon_t]$

Thus, $E[Y_t]=\phi E[ Y_{t-1}]$, $\mu=\phi \mu$

Therefore $\mu=0$. This is the only posibility for the condition to hold.

$\gamma(h)=Cov(Y_t,Y_{t+h})$

> $=E(Y_tY_{t+h})-E(Y_t)E(Y_{t+h})$ 

> $= E(Y_tY_{t+h})$

> $=E[Y_t(\phi Y_{t-1+h} +\epsilon_{t+h})]$

> $=E[\phi Y_tY_{t-1+h} +Y_t\epsilon_{t+h})]$

> $=\phi E[Y_tY_{t-1+h}] +E[Y_t\epsilon_{t+h}]$

also known that $E[Y_t\epsilon_{t+h}]=Cov(Y_t\epsilon_{t+h})=0$ because current value should not depend on future error. there is no independence between those two thing, so the cov is 0.

$\gamma(h)=E[Y_t,Y_{t+h}]=\phi E[Y_tY_{t-1+h}]$

> $=\phi^2\gamma(h-2)$

> $=\phi^3\gamma(h-3)$

> ...

> $=\phi^h\gamma(0)$

$\gamma(0)=Var[Y_t]=E[Y_t^2]-E[Y_t]$ Known $E[Y_t]=0$ 

> $=E[(\phi Y_{t-1} +\epsilon_t)^2]$

> $=E[\phi^2Y_{t-1}^2+2\phi Y_{t-1}\epsilon_t+\epsilon_t^2]$

> $=\phi^2E[Y_{t-1}^2]+2\phi E[Y_{t-1}\epsilon_t]+E[\epsilon_t^2]$

> $=\phi^2\gamma(0)+0+\sigma^2$

=> $\gamma(0)=\frac{\sigma^2}{1-\phi^2}$

Therefore $\gamma(h) = \frac{\phi^{|h|}\sigma^2}{1-\phi^2}$ for $h\in Z$ and $\rho(h)=\frac{\gamma(h)}{\gamma(0)}=\phi^{|h|}, h\in Z$.

we will have a decreasing ACF. we would expect to see a quick exponential decay.

Whereas we've calculated ACF's from specified models, in practice we observe data and calculate sample estimates of the quantities.

Given an observed time series $\{Y_t\} = \{Y_1,Y_2,Y_3...,Y_n\}$

+ Sample Mean Function: $\hat{\mu(t)} = \bar{Y} = \frac{1}{n}\sum_{t=1}^nY_t$

+ Sample Autocovariance Function: $\hat{\gamma}(h) = \frac{1}{n}\sum_{t=1}^{n-|h|}(Y_t-\bar{Y})(Y_{t+|h|}-\bar{Y})$

The bias is not significant when in a large enough time series, thus we may use n instead of n-1.

+ Sample Autocorrelation Function: $\hat{\rho}(h)=\frac{\hat{\gamma}(h)}{\hat{\gamma}(0)}$

In practice, we use the $\hat{\rho}(h)$ to determine whether an observed time series is correlated. We can determine a threshold where if $\hat{\rho}(h)$ lies beyond this threshold, The correlation at lag h is deemed to be significant. 

These threshold are calculated in the context of the fact that asymptotically $\tilde\rho(h)\sim N(0,\frac{1}{n})$

if the time series is uncorrelated. The threshold is calculated as  $\pm 1.96/\sqrt{n}$ and $\hat{\rho}(h)\notin [-1.96/\sqrt{n},1.96/\sqrt{n}]$ indicates significant lag h correlation.

## Sample ACF example

```{r}
# sample ACF for iid noise N(0,1)
X <- rnorm(100) # generating (independently) 100 realizations of N(0,1)
par(mfrow=c(2,1)) #dividing the page into 2 rows and one column
plot(X,type='l',main='iid noise') #plotting the data
acf(X,main='Sample ACF for iid noise', lag.max = 100) # plotting the acf
```

```{r}
# sample ACF for MA(1) process
X <- arima.sim(list(order = c(0,0,1), ma = 0.85), n = 200) # simulating data from an MA(1) process
par(mfrow=c(2,1)) #dividing the page into 2 rows and one column
plot(X,type='l',main='Simulated data from MA(1)') #plotting the data
acf(X,main='Sample ACF for MA(1)') # plotting the acf
```
```{r}
# sample ACF for AR(1) process
X <- arima.sim(list(order = c(1,0,0), ar = .7), n = 10000) # simulating data from an AR(1) process
par(mfrow=c(2,1)) #dividing the page into 2 rows and one column
plot(X,type='l',main='Simulated data from AR(1)') #plotting the data
acf(X,main='Sample ACF for AR(1)', lag.max = 120) # plotting the acf
arima(X,order=c(1,0,0))
#arima(X,order=c(2,0,0))
```
```{r}
# sample ACF for data with trend
a <- seq(1,100,length=200)
X <- 22-15*a+0.3*a^2+rnorm(200,500,50)
par(mfrow=c(2,1)) #dividing the page into 2 rows and one column
ts.plot(X, main = "Time Series With Significant Trend")
acf(X, main = "ACF Exhibits Seasonality + Slow Decay", lag.max = 60)
```

```{r}
# sanple ACF for data with seasonality
X <- 100*sin(1:200)+rnorm(200, 0, 0.5) 
par(mfrow=c(2,1))
ts.plot(X, main = "Time Series With Significant Seasonality")
acf(X, main = "ACF Also Exhibits Seasonality", lag.max=60)
```

```{r}
# sample ACF for data with trend and seasonal component
a <- seq(1,10,length=200)
X <- 22-15*a+a^2+5*sin(20*a)+rnorm(200,20,2)
par(mfrow=c(2,1)) #dividing the page into 2 rows and one column
ts.plot(X, main = "Time Series With Significant Trend and Seasonality")
acf(X, main = "ACF Exhibits Seasonality + Slow Decay", lag.max = 60)
```

```{r}
# sample ACF for US Accidental Deaths data (data with seasonality)
par(mfrow=c(2,1)) #dividing the page into 2 rows and one column
plot(USAccDeaths,type='l',main='# of Accidental deaths in US') #plotting the data
acf(USAccDeaths,main='Sample ACF for US accidental deaths data',lag.max=48) # plotting the acf
```

