---
title: "Nov 6 Notes"
author: "Hongdou Li"
date: "11/6/2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Recap

+ An AR(p) model is stationary iff $AR(p) = MA(\infty)$

> + This is satisfied **iff** the zeros of the AR generating function lie outside the unit circle in the complex plane.

> + This the **"stationary condition"** for AR models.

+ An MA(q) model is only useful if it has an infinite order AR representation. (ie. $MA(q)=AR(\infty)$)

> + This is satisfied **iff** the zeros of the MA generating function lie outside the unit circle in the complex plane.

> + This is the **"Invertibility condition"** for MA models.

## ARMA(p,q) Models

$\{Y_t\}$ is an autoregressive moving average of orders p and q if the following conditions are true:

$Y_t = \phi_1Y_{t-1}+\phi_2Y_{t-2}+...+\phi_pY_{t-p}+\epsilon_t+\theta_1\epsilon_{t-1}+\theta_2\epsilon_{t-2}+...+\theta_q\epsilon_{t-q}$

where $\{\epsilon_t\}\sim WN(0,\sigma^2)$ and $\phi_1,...\phi_p,\theta_1...\theta_q$ are constants.

This model can be represented in terms of generating functions as follows

\begin{center}
$Y_t - \phi_1Y_{t-1}-\phi_2Y_{t-2}-...-\phi_pY_{t-p} = \epsilon_t+\theta_1\epsilon_{t-1}+\theta_2\epsilon_{t-2}+...+\theta_q\epsilon_{t-q}$

$(1 - \phi_1B-\phi_2B^2-...-\phi_pB^p)Y_{t} = (1+\theta_1B+\theta_2B^2+...+\theta_qB^q)\epsilon_{t}$

$\phi(B)Y_t=\theta(B)\epsilon_t$
\end{center}

where $\phi(Z)=1 - \phi_1Z-\phi_2Z^2-...-\phi_pZ^p$ is the generating function of the AR component.

and $\theta(Z) = 1+\theta_1Z+\theta_2Z^2+...+\theta_qZ^q$ is the generating function of the MA component.


### Remarks:

+ An ARMA(p,q) models is not necessarily stationary, but we'd like it to be so that it could be used to model stationary time series.

+ An ARMA(p,q) model is stationary **iff** its AR component is stationary.

> + we check this by determing whether the generating function of the AR component satisfies the stationary conditions.

+ An ARMA(p,q) model is not necessarily invertible, but we'd like it to be so that we chould write $Y_t$ exclusively as a function of previous Y's

+ An ARMA(p,q) model is invertible **iff** its MA component is invertible.

> + we can check this by determing whether or not the generating function of the MA component satisfies the invertibility conditions.

## Other Comments

+ AR(p) = ARMA(p,0)

+ MA(Q) = ARMA(p,q)

+ Model selection in the context of ARMA models require us to choose p and q. We can use ACF and PACF plots to help with this:

> + ACF: q spikes (lags h $\le$ q) + (mixed) exponential decay

> + PACF: p spikes (lags h $\le$ q) + (mixed) exponential decay


## Example:

$\{Y_t\} \sim ARMA(2,3)$

$Y_t = Y_{t-1}+0.5Y_{t-2}+\epsilon_t+0.2\epsilon_{t-1}+0.7\epsilon_{t-2}$

Determine whether $\{Y_t\}$ is stationary and/or invertible

$Y_t - Y_{t-1}-0.5Y_{t-2}=\epsilon_t+0.2\epsilon_{t-1}+0.7\epsilon_{t-2}$

$(1-B-0.5B^2)Y_t = (1+0.2B+0,7B^2)\epsilon_t$

$\phi(B)Y_t=\theta(B)\epsilon_t$

where:
\begin{center}
$\phi(Z) = 1-Z-0.5Z^2$

$\phi(Z)=0$ if $Z=\frac{1\pm\sqrt{(-1)^2-4(-0.5)(1)}}{2(-0.5)}=-1\pm\sqrt{3}$
\end{center}
$Z_1 = -1-\sqrt{3} = -2.73$ and $|Z_1|=2.73$

$Z_2 = -1+\sqrt{3} = 0.73$ and $|Z_2|=0.73$

since $|Z_2| < 1$ this root lies inside the unit circle and so the ARMA model is not stationary.
\begin{center}
$\theta(Z) = 1+0.2Z+0,7Z^2$

$\theta(Z)=0$ if $Z=\frac{-0.2\pm\sqrt{(0.2)^2-4(-0.7)(1)}}{2(-0.7)}=\frac{-2\pm\sqrt{-2.76}}{1.4}=\frac{-2\pm\sqrt{2.76}i}{1.4}$
\end{center}
$Z_1 = -0.14-1.19i$ and $|Z_1|= \sqrt{(-0.14)^2+(-1.19)^2}=1.198$

$Z_2 = -0.14+1.19i$ and $|Z_2|=1.198$

since $|Z_1|=|Z_2| > 1$ this root lies outside the unit circle and so the ARMA model is invertible.


just identify it is an ARMA model, choosing the exact p and q can be very difficult.
```{r}
#data.sim <- arima.sim(n=10000, list(ar=c(0.3,0.65),ma=c(-0.5),sd=sqrt(1)))
#auto.arima(data.sim)
```

## Estimating ARMA(p,q) models

Goal: to estimate $\phi_1,\phi_2,...,\phi_p,\theta_1,\theta_2,...\theta_q,\sigma$ in the context of an ARMA model: 

$\phi(B)Y_t=\theta(B)\epsilon_t$

Those parameters are estimated with observed data $\{y_1,y_2,...y_n\}$

several methods of estimation exist, but we'll focus on :

(1) Maximum likelihood estimation

(2) Least Squares Estimation


